---
title: "SQL and R"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Setting up the connection

#### Setup

Our first step is to prepare RStudio to set up a connection to a mySQL database.  

For this, we will need two things:

1. Download a `SQLIte` database to your machine, so we can practice with it:
https://cdn.sqlitetutorial.net/wp-content/uploads/2018/03/chinook.zip

2. Install some SQL-related R packages:

```{r}
#install.packages(c("RSQLite", "DBI", "dbplyr", "obdc"))
```

Next we'll load libraries (of course):

```{r}
library(RSQLite)
library(DBI)
library(odbc)
library(dbplyr)
library(tidyverse)
```
Last we'll verify that the connections are working properly

```{r}
sort(unique(odbcListDrivers()[[1]]))
```


#### Connect to database

Now we set up the connection:

```{r}
### You may need to replace "chinook.db" with the full path to this file on your machine.
con <- dbConnect(SQLite(), "chinook.db")
```


## Getting a Dataset

Normally, you'd explore the data by looking at the object with `summary()` or
similar.  However, our object `con` isn't a **Data Frame**, but instead, an
object of type *SQLiteConnection*.

```{r}
con
summary(con)
```

Let's find out what datasets are in this database:

```{r}
src_dbi(con)
```


To look a particular dataset in the database, we can use the `tbl()` function from `dbplyr`.

```{r}
tbl(con, "customers")
```

Let's save this table as an object in R:

```{r}
cust <- tbl(con, "customers")

cust
```

Notice that although `emp` appears to be a data frame, it is actually just a
"pointer" to where the `employees` dataset lives in the database, as well as
some information about its row and column names.

```{r}
str(cust)
```

We'll never be able to use our familiar "shortcut" functions, like `summary()`,
because these operate on the full dataset.  We are trying to avoid pulling the
full dataset into R, presumably because it is too large.

```{r}
summary(cust)
```


## Data Manipulation

Fortunately, the `dbplyr` packages implements most of the main `dplyr` operations.

Instead of running these operations in R, `dbplyr` "translates" them into SQL
code, and then sends that to the database.

For example,

```{r}
cust %>%
  select(LastName, FirstName, CustomerId) %>%
  filter(CustomerId %% 2 == 0) %>%
  show_query()
```

Once again, this is **never** returning the full data frame to R:

```{r}
cust %>%
  select(LastName, FirstName) %>%
  nrow()
```

It is simply translating R to SQL:

```{r}
cust %>%
  select(LastName, FirstName) %>%
  show_query()
```


The advantage is that we can do simple operations on very large data using SQL
**much** faster than in R.  (In fact, many datasets would be impossible even
to read into R.)

Other than that "hidden" advantage, your dplyr code looks the same as it normally would:

```{r}
cust %>%
  count(Country)
  
```

```{r}
cust %>%
  group_by(country) %>%
  summarize(n())
```

## Other Operations

#### Collecting

There may be times that you do want to pull a portion of the dataset into R.
You can do this with `collect()`

Be careful, though!  Make sure what you are loading in is of reasonable size!

```{r}
cust_bra <- cust %>%
  filter(country == "Brazil") %>%
  collect()

cust_bra
```

Here the `filter` step is done in SQL, but then `cust_bra` becomes simply a
data frame in R.

Any further operations you do happen in R, not SQL:

```{r}
cust_bra %>%
  select(FirstName)
```

```{r, error = TRUE}
cust_bra %>%
  select(FirstName) %>%
  show_query()
```

#### Writing to Databases

You can also use your connection to add to a database:

```{r}
dbWriteTable(con, "brazilian_customers", cust_bra)
```

```{r}
src_dbi(con)
```

#### Statistics

The tradeoff of SQL's speed is its flexibility.  Many fancy statistical methods
aren't implemented in SQL/dbplyr.

However, a lot can be done!

```{r}
emp %>%
  lm(SupportRepId ~ Country, data = .)
```

## Connecting to a real database

```{r}
con <- dbConnect(RMySQL::MySQL(), 
                 user='kbodwin', 
                 password='somepassword', 
                 dbname='vineyard_diversity',
                 host='vineyard-project.cnmgssyjqpsv.us-west-1.rds.amazonaws.com',
                 port=3306)
```

