---
title: "Projects, Directories, Data Files"
format: 
  revealjs:
    theme: sky
execute:
  include: true
  echo: true
  eval: true
---

## Principles of Reproducibility

The idea:  You should be able to send your project to someone else, and they can jump in and start working right away.

1. Files are organized and well-named.

2. References to data and code work for everyone.

3. Package dependency is clear.

4. Code will run the same every time, even if data values change.

5. Analysis process is well-explained and easy to read.

```{r}
#| include: false
library(tidyverse)
```


## Good practice

- Organize your folders carefully, and name them meaningfully:  
    + `/User/kbodwin/Dropbox/Workshops/` rather than `Desktop/stuff/`
    
- Use **R Projects** liberally - put one in the "main" folder for each project

    + I have a project for every class I teach.
    + I have a project for every data analysis I do.
    + I **never** open RStudio, except by clicking on a project.

## Bad practice

If you put something like this at the top of your .Rmd file, I will set your computer on fire:

```{r}
#| eval: false
#| echo: true
setwd("/User/reginageorge/Desktop/R_Class/Lab_1/")
```

- Setting working directory by hand = BAD!  

- That directory is specific to you!

- (Quarto files may not render this code.)


# Locating your data files
 
## read_csv()

- The *tidyverse* way to read data is with `read_csv()`.

- To tell this function where to find the data, you can give a **url** to the dataset, or a **path**.

```{r}
#| eval: false
#| echo: true
hiphop <- read_csv("https://www.dropbox.com/s/1qqyshx5ikt9zoc/hiphop.csv?dl=1")
## This will work for everyone!

```

. . .

```{r}
#| eval: false
#| echo: true

hiphop <- read_csv("/Users/kbodwin/Dropbox/Data/hiphop.csv")
## This will work only on my computer

```

. . .

```{r}
#| eval: false
#| echo: true

hiphop <- read_csv("hiphop.csv")
## This will work if the file is in the same directory as the code (i.e., the project directory)
```

. . .

```{r}
#| eval: false
#| echo: true

hiphop <- read_csv("./Data/hiphop.csv")
## This will work if a folder called "Data" is in the same directory as the code.
```


## The **here** package

A great solution to consistency in paths is the `here` package:

> install.packages("here")

This package thinks that *"here"* means *"the directory of the closest project"*.

```{r}
#| message: true
here::dr_here()
```


## The **here** package

These slides live in a subfolder called `Slides`.

The *data* is not in this folder!

. . .

If I wanted to access a dataset from my main Data folder, I'd have to use a relative file path:

```{r}
#| eval: false
hiphop <- read_csv("../../Data/hiphop.csv")
```

. . .

If I ever reshuffle my folder structure, I'm in trouble!


## The **here** package

Instead of trying to reference data relative to **this** file, let's reference it relative to the **Project** location, which is unlikely to change.

```{r}
#| eval: false
here::here("Data", "hiphop.csv")
```

. . .

```{r}
#| eval: false
hiphop <- read_csv(here::here("Data", "hiphop.csv"))
```

This will **adapt** to each person's path!



# Reading in data



## Common types of data files

- `.csv` :  "Comma-separated"
```
Name, Age
Bob, 49
Joe, 40
```


## Common types of data files

- `.txt`:  Plain text
    + Could be just text
    + Could be comma-separated data
    + Could be tab-separated, bar-separated, etc.
    + Need to let R know what to look for

- `.xls, .xlsx`:  Microsoft Excel Spreadsheet
    + Common approach: save as `.csv`
    + Nicer approach:  `readxl` package



## Separators

- Tell R what symbol separates data observations with `delim = `

```{r}
#| eval: false
hiphop <- read_csv("hiphop.csv", delim = ",")
hiphop <- read.table("hiphop.csv", delim = "\t")  
## if tabs between data
```


## Try it together!

Your `Data` folder contains several datasets containing the names and ages of five men.  The datasets are stored as different file types.  

1. Read in the dataset `ages.csv`.
2. Read in the dataset `ages_tab.txt`
3. Read in the dataset `ages_mystery.txt`
4. Find a way to use `read_csv` to read `ages.csv` with the variable "Name" as a factor and "Age" as a character.
5. *(Optional)* Read in the dataset `ages.xlsx`


## Your Turn

Open the Activity `02-Explore_Data.qmd` to see a few summaries of this dataset.